Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-04 files/Chapter2/generalvibrationmodel.tex
presentation
Chapter 4 + 5 review

\subsubsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

@@

\subsubsection{Comparison of models}

In this dissertation, the objective is to compare different linear beam and plate models for use in applications. The first of the comparisons are in Section \ref{sec:SP06}. This section is a discussion of the article \cite{SP06} comparing a Timoshenko beam model and a three dimensional beam model to empirical results. The authors of this article use the natural frequencies (equivalently the eigenvalues) to compare the different models.\\

@@

These comparisons are made in Chapter 4 and 6. Being able to express the solutions of the models as a formal series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models. This is explained in detail in \cite{CVV18}.

The validity of the formal series solution with initial conditions is required to investigate the vibration of the models. This ensured that if the models are disturbed in the exact same way, the formal series solution remains valid, and a comparison can be made.

>>>>

\subsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

@@

\subsection{Comparison of models}

In this dissertation, the objective is to compare different linear beam and plate models for use in applications. The first of the comparisons are in Section \ref{sec:SP06}. This section is a discussion of the article \cite{SP06} comparing a Timoshenko beam model and a three dimensional beam model to empirical results. The authors of this article use the natural frequencies (equivalently the eigenvalues) to compare the different models.

@@

These comparisons are made in Chapter 4 and 6. Being able to express the solutions as a valid series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models. This is explained in detail in \cite{CVV18}.

In \cite{CVV18}, the authors consider a beam model and wave equation model for a vibrating string. Suppose that for the same initial conditions, $u_b$ and $u_w$ are the exact solution. A comparison of the exact solutions are not possible, but the partial sums can be compared if the bounds for the errors $u_b - u^N_b$ and $u_w - u^N_w$ can be guaranteed.

### 2023-11-04 files/Chapter4/CantileverTimo.tex
Chapter 4 + 5 review

\documentclass[../../main.tex]{subfiles}
\begin{document}
\noindent
\section{Cantilever Timoshenko beam with tip-body}
\sout{This section discusses the technical report [DBVXX] on a cantilever Timoshenko beam with a tip body and elastic boundary interfaces. Most papers on a similar topic focus on the analysis of vibration modes, stabilization and optimal control of vibrations. The goal of [DBVXX] is to realistically model the interface between the endpoint of a beam and the attached tip body. Few publications consider this modelling.}\\

\sout{In this section, the finite element analysis and some numerical simulations for the model are presented. These numerical simulations are to show that the model behaves in a realistic manner.}\\

This section discusses a technical report [DBVXX]. In this technical report the authors models a cantilever Timoshenko beam with a tip body and elastic boundary conditions. This model can be seen as a motivation of the importance of investigating the validity of the Timoshenko beam theory.\\

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\noindent
\section{Cantilever Timoshenko beam with tip-body}

\textcolor{red}{MOONTLIK UIT HAAL.}

This section discusses a technical report [DBVXX]. In this technical report the authors models a cantilever Timoshenko beam with a tip body and elastic boundary conditions. This model can be seen as a motivation of the importance of investigating the validity of the Timoshenko beam theory.

### 2023-11-04 files/Chapter5/2D_FEM.tex
Chapter 4 + 5 review

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the models of this dissertation, presented in chapter 1. The goal of this chapter is to obtain numerical formulas that can be used to calculate the eigenvalues and eigenvectors for the models. This is in preperation of chapter 6, where the eigenvalues and eigenvectors of the models are compared.

In chapter 4, a method was discussed to obtain the eigenvalues and eigenvectors for the Timoshenko beam model. Therefore it is not neccesary to apply the Finite Element Method on the Timoshenko model to obtain the eigenvalues and eigenvectors. The Finite Element Mehtod is applied to the two-dimensional and three-dimensional models.

@@

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. Recall this is Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

See chapter 2 for the existence theory.

\subsection{Galerkin approximation}\label{2d_FEM_G}
To be able to apply the Finite Element Method to Problem 2D-1V, the body $\Omega$ needs to be discretised. This is done by dividing the body $\Omega$ into discrete shapes, called elements. There are various types of shapes of elements that can be used. Since the body $\omega$ is has a rectangular cross-section, rectangular are the simplist elements to use.

@@

\subsection{Eigenvalue problem}\label{2dFEM_EP}
For the eigenvalue problem, assume that $M_{f}Q^I = 0$ so that 
\begin{eqnarray}
		M\ddot{\bar{u}} & = & K\bar{u}.\label{eq:2DFEM:M2}
\end{eqnarray}
It is known that a system of ordinary differential equations have a general solution of the form $e^{rt}$. Suppose that $\bar{w} = e^{\lambda t} \bar{u}$ is a solution for \eqref{eq:2DFEM:M2}. In this solution, $\lambda$ is an eigenvalue and $\bar{u}$ a corresponding eigenfunction. Substitution into \eqref{eq:2DFEM:M2} results in
\begin{eqnarray*}
	M\lambda e^{\lambda t}\bar{u} & = & Ke^{\lambda t}\bar{u}.
\end{eqnarray*}
Since $e^{\lambda t} > 0$ for all values of $\lambda t$, we can cancel it from both sides of the equation and formulate the eigenvalue problem Problem 2D-1E.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the models of this dissertation, presented in Chapter 1. The goal of this chapter is to obtain an algorithm that can be used to calculate the eigenvalues and eigenvectors for the models. This is in preparation of chapter 6, where the eigenvalues and eigenvectors of the models are compared.

In chapter 4, a method was discussed to obtain the eigenvalues and eigenvectors for the Timoshenko beam model. Therefore it is not necessary to apply the Finite Element Method on the Timoshenko model to obtain the eigenvalues and eigenvectors. The Finite Element Method is applied to the two-dimensional and three-dimensional models.

@@

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. This is Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

See Chapter 2 for the existence theory.

\subsection{Galerkin approximation}\label{2d_FEM_G}
To be able to apply the Finite Element Method to Problem 2D-1V, the body $\Omega$ needs to be discretised. This is done by dividing the body $\Omega$ into discrete shapes, called elements. There are various types of shapes of elements that can be used. Since the body $\omega$ is has a rectangular cross-section, rectangular elements are the simplest elements to use.

@@

\subsection{Eigenvalue problem}\label{2dFEM_EP}
For the eigenvalue problem, assume that there is no external force, $M_{f}Q^I = 0$, so that 
\begin{eqnarray}
		M\ddot{\bar{u}} & = & K\bar{u}.\label{eq:2DFEM:M2}
\end{eqnarray}
It is known that a system of ordinary differential equations has a general solution of the form $e^{rt}$. Suppose that $\bar{w} = e^{\lambda t} \bar{u}$ is a solution for \eqref{eq:2DFEM:M2}. In this solution, $\lambda$ is an eigenvalue and $\bar{u}$ a corresponding eigenfunction. Substitution into \eqref{eq:2DFEM:M2} results in
\begin{eqnarray*}
	M\lambda e^{\lambda t}\bar{u} & = & Ke^{\lambda t}\bar{u}.
\end{eqnarray*}
Since $e^{\lambda t} > 0$ for all values of $\lambda t$, we can cancel it from both sides of the equation and formulate the eigenvalue problem Problem 2D-1E.

### 2023-11-04 files/Chapter5/3D_FEM.tex
Chapter 4 + 5 review

In section \ref{ssec:3D_Model:VariationalForm}, the variational problem for the three-dimensional cantilever model is defined by Problem 3D-1V. This general form can now be rewritten as in a model specific form using the reference configuration. For convenience, some of the results from section \ref{sec:3D_Mode} are repeated here.

@@

With the test function space
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C(\Omega) \ | \ \phi = 0 \ \textrm{ on } \ \Sigma_0 \right\}.
\end{eqnarray*}

@@

See chapter 2 for the existence theory.

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. This is model, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as brick shaped elements as described by \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.\

@@

For the this three-dimensional model, tri-cubix basis functions are used. These basis functions are an extension of the Hermite cubic basis functions from the two-dimensional model. Although this is more complex than using tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.\\

For this dissertation, we use tri-cubic functions $\phi_i$. For a simplification, tri-linear functions can also be used instead. For easier notation, denote each of the basis functions in the set by $\delta_i$.\\

Similar to \ref{2d_FEM_G}, denote the admissible basis functions by $\delta_j$, where each $\delta_j$ is a unique element of $B$ satisfying all the conditions of the test function space $T(\Omega)$. The admissible basis functions are numbered and expressed as the set $A = \left\{\delta_1, \delta_2, ..., \delta_k \right\}$ for some $k \leq 3n$.\\

>>>>

In section \ref{ssec:3D_Model:VariationalForm}, the variational problem for the three-dimensional cantilever model is defined by Problem 3D-1V. This general form can now be rewritten as in a model specific form using the reference configuration. For convenience, some of the results from section \ref{sec:3D_Model} are repeated here.

@@

With the test function space
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C(\Omega) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}

@@

See Chapter 2 for the existence theory.

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. For this is three-dimensional body, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as `brick-shaped' elements and is described in \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.

@@

For the this three-dimensional model, tri-cubix basis functions are used. These basis functions are an extension of the Hermite cubic basis functions from the two-dimensional model. Although this is more complex than using tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.

Recall that the admissible basis functions, are all the basis functions that satisfies all the conditions of the test function space $T(\Omega)$. Denote the admissible basis functions by $\delta_j$, with each $\delta_j$ a unique element of $B$. The set of admissible basis functions can then be expressed as $A = \left\{\delta_1, \delta_2,... , \delta_k\right\}$ with $k \leq 3n$.

### 2023-11-04 files/Chapter5/PlateModel.tex
v2 chapter 7.2, started 7.3
Chapter 4 + 5 review

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Cantilever plate model}\label{sec:FEM:Plate}
Consider a rectangular cantilever Reissner-Mindlin plate.

\subsubsection{Reference configuration for a rectangular plate}
Let $\left\{e_1,e_2\right\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body as $\Omega \in R^2$ with the reference point $(0,0)$. For a rectangular plate,
\begin{eqnarray*}
	\Omega = \left \{ x \in R_2 \ | \ 0 \leq x_1 \leq 1, \ 0 \leq x_2 \leq b \right \}.
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of plate. The boundary $\partial \Omega$ can be divided into four distinct lines.

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma_0:& \quad x_1 &= 0\\
		\Sigma_1:& \quad x_1 &= 1
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_0:& \quad x_2 &= 0\\
		\Gamma_1:& \quad x_2 &= b
	\end{eqnarray*}
\end{minipage}\\

\textbf{Remark} The mathematical definition of $R_2$ is the same as the two-dimensional beam. Practically, it is not the same. The plate model is a horisontal two-dimensional plane, and the two-dimensional beam is a vertical two-dimensional plane.

\subsubsection{Cantilever plate model}
Consider a rectangular plate clamped rigidly to a surface at the edge $\Sigma_0$. The plate is free hanging on the remaining edges and there is not external force acting on the plate. The plate is be modelled using the Reissner-Mindlin plate theory that is presented in section \ref{ssec:P_Model:ModelProblems} as Problem P-1.\\

@@

	\end{tikzpicture}
	\caption{Two-dimensional Cantilever Plate}
\end{figure}
\FloatBarrier

@@

\subsection{Weak variational form}
Similar to Section 2.1, the weak variational form for Problem P-1 can be derived from Problem P-1V.
\subsubsection{Bilinear forms}
From the bilinear form, we have
\begin{eqnarray*}
	b(f_2,g_2) & = & \frac{1}{\beta(1-\nu^2)}\iint_{\Omega} (\partial_1f_{2,1} + \nu\partial_2f_{2,2})\partial_1g_{2,1}+ (\partial_2f_{2,2} + \nu\partial_1f_{2,1})\partial_2g_{2,2} \ dA,\\
	& + & \frac{1}{2\beta(1+\nu)}\iint_{\Omega} (\partial_1f_{2,2}+\partial_2f_{2,1})(\partial_1g_{2,2}+\partial_2 g_{2,1}) \ dA
\end{eqnarray*}

@@

\subsubsection{Problem Plate-1W}
Find a function $u$ such that for all $t>0$, $u(t) \in V$, $u'(t) \in V$ and $u''(t) \in W$, satisfying the following equation
\begin{eqnarray}
	c(u''(t),v) + b^*(u(t),v) & = & (f(t),v)_{\Omega} \ \ \ \textrm{ for each } v \in V,
\end{eqnarray} with $u(0)= u_0 = \langle w_0, \boldsymbol\psi_0 \rangle$, and $u'(0)= u_1 = \langle w_1, \boldsymbol\psi_1 \rangle$.\\

\subsubsection{Existence and uniqueness}
\textcolor{red}{XXXXXXXXXXXXXXXXXXXXX}
Section 2.2 provides the necessary theory for the existence and uniqueness of solutions for second order hyperbolic equations. The assumptions \textbf{A1}, \textbf{A2} and \textbf{A4} are similar to Section 2.3. So it is only required to show that $b$ is an inner product for $V$ and that there exists a positive number $C_V$ such that
\begin{eqnarray*}
	\Vert v\Vert_{W} \leq C_{V}\Vert v\Vert_{V}
\end{eqnarray*} for each $v \in V$.\\

However, this is not as simple as the example in Section 2.4. ****Reference 'n article wat na dit kyk.***

\subsection{Galerkin approximation}
Divide the reference configuration $\Omega$ into a rectangular grid of elements, such that there are $n = n_1 \times n_2$ nodes.\\

For the plate model, two different definitions for basis functions are required.\\

First consider a set of $n$-dimensional linear independent basis functions, defined by the set
\begin{eqnarray*}
	\left\{\delta^*_1, \delta^*_2, ..., \delta^*_n\right\}.
\end{eqnarray*}
In this set, each $\delta^*$ is a cubic function. The subset of functions that satisfies the conditions of the test function space $T_1(\bar{\Omega})$ are the admissible basis functions for $T_1(\bar{\Omega})$. These admissible basis functions can be renumbered and expressed as the set $\left\{\delta^*_1, \delta^*_2, ..., \delta^*_{k_1}\right\}$ for some $k_1 \leq 2n$.\\

Next consider a set of $n$-dimensional linear independent basis functions, defined by the set
\begin{eqnarray*}
	\left\{\langle\phi_1, 0\rangle, \langle\phi_2, 0\rangle,...,\langle\phi_{n}, 0 \rangle,\langle 0,\phi_1\rangle,\langle 0 ,\phi_2\rangle,...,\langle 0,\phi_{n}\rangle \right\}.
\end{eqnarray*} These functions $\phi_j$ are cubic basis functions. Denote each element in this set by $\delta_i$. The subset of functions that satisfies the conditions of the test function space $T_2(\bar{\Omega})$ are the admissible basis functions for $T_2(\bar{\Omega})$. These admissible basis functions can be renumbered and expressed as the set $\left\{\delta_1, \delta_2, ..., \delta_{k_2}\right\}$ for some $k_2 \leq 2n$.\\

Define the two spaces
\begin{eqnarray*}
	S^h_1 & = & \textrm{span}\left(\left\{\delta^*_i \ | \ i = 1,2,...,k_1 \right\} \right),\\
	S^h_2 & = & \textrm{span}\left(\left\{\delta_i \ | \ i = 1,2,...,k_2 \right\} \right).
\end{eqnarray*}
For each function $u^h \in S_1^h$ and each function $\psi^h \in S_2^h$, $u^h$ can be expressed as
\begin{eqnarray*}
	w^h = \sum_{i = 1}^{k} w_i(t) \delta^*_{i}(x)
\end{eqnarray*} and $\psi^h$ can be expressed as
\begin{eqnarray*}
	\psi^h = \sum_{i = 1}^{k} \psi_i(t) \delta_{i}(x)
\end{eqnarray*}

Problem P1-V can now be rewritten into a Galerkin Approximation, denoted by Problem P-1G.

@@

\subsection{System of ordinary differential equations}
The standard FEM matrices $K_{11}$ to $K_{22}$ and $M_F$ where presented in section \ref{ssec:2DFEM:DE}. In addition, the following standard FEM matrices are also required.

@@

for $i,j = 1,2,...,p$.\\

Define the following matrices:\\
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{K11} & = & hK_{11}-hK_{22}\\
		\mathbf{K12} & = & hL_1\\
		\mathbf{K13} & = & hL_2\\
		\mathbf{K21} & = & hL_1^T\\
		\mathbf{K22} & = & AK_{11}+BK_{22}+h\mathbf{M}
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{0.8\linewidth}
	\begin{eqnarray*}
		\mathbf{K23} & = & A\nu K_{12}+BK_{21}\\
		\mathbf{K31} & = & hL_2^T\\
		\mathbf{K32} & = & A\nu K_{21}+BK_{12}\\
		\mathbf{K33} & = & AK_{22}+BK_{11}+h\mathbf{M}
	\end{eqnarray*}
\end{minipage}\\

with $A = \frac{1}{\beta(1-\nu^2)}$ and $B = \frac{1}{2\beta(1+\nu)}$.\\

@@

Using \eqref{eq:PFEM:K+M} and \eqref{eq:PFEM:M}, Problem P-1G is rewritten as a system of ordinary differential equations. This system is referred to as Problem P-1ODE
\subsubsection{Problem P-1ODE}
Find function $\bar{u} \in S_1^h\times S_2^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I \label{Plate_FEM_M}
\end{eqnarray}
With $\bar{u}$ in the form $\bar{u} = \langle w, \partial_x w, \partial_y w, \partial_{xy} w, \psi_1, \partial_x \psi_1, \partial_y \psi_1, \partial_{xy} \psi_1, \psi_2, \partial_x \psi_2, \partial_y \psi_2, \partial_{xy} \psi_2 \rangle$.

@@

The equation \eqref{Plate_FEM_M} is the same form as in section \ref{2dFEM_EP} for the two-dimensional elastic body. Therefore the derivation of the eigenvalue problem is identical. Denote the eigenvalue problem for Problem P-1 by Problem P-1E.
\subsubsection{Problem P-1E}
Find a vector function $u$ and a number $\lambda$ such that
\begin{eqnarray}
	\lambda M{u} & = & K{u}.
\end{eqnarray}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Cantilever plate model}\label{sec:FEM:Plate}
Consider a rectangular cantilever Reissner-Mindlin plate model with a rectangular cross-section.

\subsubsection{Reference configuration for a rectangular plate}
Let $\left\{e_1,e_2\right\}$ be a right-handed orthonormal basis for $R^2$. Although this basis seems identical to the two-dimensional beam in section \ref{sssec:2D_Model:RefConf}, it is fact different. It would be more appropriate to use $e_1$ and $e_3$ for this plate model. However the use of $e_1$ and $e_2$ is kept to reiterate that this is in fact a two-dimensional model.

Denote the elastic body as $\Omega \in R^2$ with the reference point $(0,0)$. For a rectangular plate,
\begin{eqnarray*}
	\Omega = \left \{ x \in R_2 \ | \ 0 \leq x_1 \leq 1, \ 0 \leq x_2 \leq b \right \}.
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of plate. The boundary $\partial \Omega$ can be divided into four distinct lines.

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma:& \quad x_1 &= 0\\
		\Gamma_3:& \quad x_1 &= 1
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_0:& \quad x_2 &= 0\\
		\Gamma_1:& \quad x_2 &= b
	\end{eqnarray*}
\end{minipage}

Similar to the two and three-dimensional cantilever models, this notation implies that the plate is clamped at $\Sigma$ and free hanging at $\Gamma$.

\subsubsection{Cantilever plate model}
Consider a rectangular Reissner-Mindlin plate clamped rigidly to a surface at $\Sigma$ and free hanging on the remaining edges. This plate model is presented in section \ref{ssec:P_Model:ModelProblems} as Problem P-1.

@@

	\end{tikzpicture}
	\caption{Two-dimensional cantilever Reissner-Mindlin plate}
\end{figure}
\FloatBarrier

@@

\subsection{Weak variational form}
Similar to Section 2.1, the weak variational form for Problem P-1 can be derived from Problem P-1V.

\subsubsection{Bilinear forms}
From the bilinear form, we have
\begin{eqnarray*}
	b(f_2,g_2) & = & \frac{1}{\beta(1-\nu^2)}\iint_{\Omega} (\partial_1f_{2,1} + \nu\partial_2f_{2,2})\partial_1g_{2,1}+ (\partial_2f_{2,2} + \nu\partial_1f_{2,1})\partial_2g_{2,2} \ dA,\\
	& + & \frac{1}{2\beta(1+\nu)}\iint_{\Omega} (\partial_1f_{2,2}+\partial_2f_{2,1})(\partial_1g_{2,2}+\partial_2 g_{2,1}) \ dA
\end{eqnarray*}

@@

\subsubsection{Problem Plate-1W}
Find a function $u$ such that for all $t>0$, $u(t) \in V$, $u'(t) \in V$ and $u''(t) \in W$, satisfying the following equation
\begin{eqnarray}
	c(u''(t),v) + b^*(u(t),v) & = & (f(t),v)_{\Omega} \ \ \ \textrm{ for each } v \in V,
\end{eqnarray} with $u(0)= u_0 = \langle w_0, \boldsymbol\psi_0 \rangle$, and $u'(0)= u_1 = \langle w_1, \boldsymbol\psi_1 \rangle$.

See Chapter 2 for the existence theory.

\subsection{Galerkin approximation}
To discretise the body $\Omega$, the same shapes as in section \ref{ssec:2D_Model:Galerkin} are used.

Divide the reference configuration $\Omega$ into a rectangular grid of elements, such that there are $n = n_1 \times n_2$ nodes.

Define a set of $n$-dimensional linear independent basis functions. The basis functions can be defined by the set $$B = \left\{\langle\phi_1, 0\rangle, \langle\phi_2, 0\rangle,...,\langle\phi_{n}, 0 \rangle,\langle 0,\phi_1\rangle,\langle 0 ,\phi_2\rangle,...,\langle 0,\phi_{n}\rangle \right\}.$$

These basis functions are bi-cubic basis functions.

Since there are two test function spaces, $T_1(\Omega)$ and $T_2(\Omega)$, two different sets of admissible basis functions are required. Denote the admissible basis functions for $T_1(\Omega)$ by $\delta^1_j$ where each $\delta^1_j$ is a unique element of B. The set of admissible basis functions that satisfies $T_1(\Omega)$ can be defined as $A_1 = \left\{\delta^1_1, \delta^1_2,..., \delta^1_{k_1} \right\}$ for a $k_1 \leq 2n$. Similarly for $T_2(\Omega)$, the set of admissible basis functions can be defined as $A_2 = \left\{\delta^2_1, \delta^2_2,..., \delta^2_{k_2} \right\}$ for a $k_2 \leq 2n$.

Define the two spaces
\begin{eqnarray*}
	S^h_1 & = & \textrm{span}\left(\left\{\delta^1_i \ | \ i = 1,2,...,k_1 \right\} \right),\\
	S^h_2 & = & \textrm{span}\left(\left\{\delta^2_i \ | \ i = 1,2,...,k_2 \right\} \right).
\end{eqnarray*}
For each function $u^h \in S_1^h$ and each function $\psi^h \in S_2^h$, $u^h$ can be expressed as
\begin{eqnarray*}
	w^h = \sum_{i = 1}^{k} w_i(t) \delta^*_{i}(x)
\end{eqnarray*} and $\psi^h$ can be expressed as
\begin{eqnarray*}
	\psi^h = \sum_{i = 1}^{k} \psi_i(t) \delta_{i}(x)
\end{eqnarray*}

Substitution of $u^h$ and $\psi^h$ into Problem P-1V, results in the following Galerkin approximation, denoted by Problem P-1G.

@@

\subsection{System of ordinary differential equations}
The standard FEM matrices $K_{11}$, $K_{12}$, $K_{21}$, $K_{22}$ and $M_F$ where presented in section \ref{ssec:2DFEM:DE}. Even though the definition of $R_2$ is different in section \ref{sec:FEM:2D}, the definition of the matrices are the same and are not repeated here.

In addition to these matrices, the following standard FEM matrices are also required.

@@

for $i,j = 1,2,...,p$.

Define the following matrices:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{K11} & = & hK_{11}-hK_{22}\\
		\mathbf{K12} & = & hL_1\\
		\mathbf{K13} & = & hL_2\\
		\mathbf{K21} & = & hL_1^T\\
		\mathbf{K22} & = & AK_{11}+BK_{22}+h\mathbf{M}
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{0.8\linewidth}
	\begin{eqnarray*}
		\mathbf{K23} & = & A\nu K_{12}+BK_{21}\\
		\mathbf{K31} & = & hL_2^T\\
		\mathbf{K32} & = & A\nu K_{21}+BK_{12}\\
		\mathbf{K33} & = & AK_{22}+BK_{11}+h\mathbf{M}
	\end{eqnarray*}
\end{minipage}\\

with $A = \frac{1}{\beta(1-\nu^2)}$ and $B = \frac{1}{2\beta(1+\nu)}$.

@@

Using \eqref{eq:PFEM:K+M} and \eqref{eq:PFEM:M}, Problem P-1G is rewritten as a system of ordinary differential equations. This system is referred to as Problem P-1ODE
\subsubsection{Problem P-1ODE}
Find function $\bar{u} \in S_1^h\times S_2^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I \label{Plate_FEM_M}
\end{eqnarray}
With $\bar{u}$ in the form $\bar{u} = \langle w, \partial_1 w, \partial_2 w, \partial_{12} w, \psi_1, \partial_1 \psi_1, \partial_2 \psi_1, \partial_{12} \psi_1, \psi_2, \partial_1 \psi_2, \partial_2 \psi_2, \partial_{12} \psi_2 \rangle$.

@@

The equation \eqref{Plate_FEM_M} is the same form as in section \ref{2dFEM_EP} for the two-dimensional elastic body. Therefore the derivation of the eigenvalue problem is identical. Denote the eigenvalue problem for Problem P-1 by Problem P-1E.
\subsubsection{Problem P-1E}
Find a vector function $\bar{u}$ and a number $\lambda$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}
